### 2016-03-14 ch05-initvals/initial-values.tex
Some work on conclusion
Some legends pulled into plots

In this chapter we assume that the output signal is disturbed by random noise, resulting in noisy error-prone data.
For the simulations, white noise is used ($H_0 = 1$ and $e(t) = v(t)$).
It is also possible to apply the estimation procedure in this chapter to a system that is disturbed by colored noise, as will be demonstrated in \secref{se:ExpMeas}.

@@

With reference to equation \eqref{lpmtd1}, the relation between the noiseless input and the output signals ($v(t)= 0$) is assumed to be of the form
\begin{equation}
    A(q^{-1}) y_0(t) = B(q^{-1}) u_0(t) \quad \forall t \in n \Ts{}  \text{with}  n \in \IntegerNumbers
\label{ABpolys}
\end{equation}
where $A$ and $B$ are polynomials in $q^{-1}$. 
Thus, it follows from equation \eqref{lpmtd1} that
\begin{equation}
    \true{G}(q^{-1})= \frac{B(q^{-1})}{A(q^{-1})}
    \text{.}
\end{equation}

From~\citet[Section 6.3.2.]{Pintelon2012}, in the frequency domain, and for the \glspl{DFT} of the windowed signals, equation \eqref{ABpolys} is of the form
\begin{align}
A(e^{-j\omega_k})Y_0(k) = B(e^{-j\omega_k})U_0(k) + I(e^{-j\omega_k})
\label{DFTspectra}
\end{align}
with $q^{-1}$ sampled in  $q_k^{-1} = e^{-j\omega_k}$, where $\omega_k = \frac{2\pi k}{N}$ are the \gls{DFT} frequencies, and $I$ is a polynomial of order $\max(N_a,N_b) - 1$, which depends on the difference between the initial and end conditions of the filters $\true{G}$ and $\true{H}$ during the measurement record.

@@

The maximum likelihood estimate of the parameter vector $\theta$  containing the coefficients of the $A$, $B$ and $I$  polynomials is obtained by solving the optimization problem
\begin{equation}
  \hat{\theta} = \arg\min_\theta V(\theta) \text{.}
\end{equation}
Since the noise $v(t)$ is assumed to be Gaussian, $V(\theta)$ accords to the weighted least squares cost function~\citep[Section 9.11]{Pintelon2012}:
\begin{equation}\label{eq:MLEcf}
V(\theta) = \sum_{k=1}^{N/2-1}\frac{|\varepsilon(k,\theta)|^2}{\sigma_\varepsilon^2(k,\theta)}
\end{equation}
for $N$ measurements and when $\varepsilon$ denotes the error in equation \eqref{DFTspectra},  \emph{viz}.:
\begin{align}
\varepsilon(k,\theta) = A(k,\theta)Y(k) - B(k,\theta)U(k) - I(k,\theta)\text{.}
\end{align}
The variance of the error is of the form:
\begin{equation}\label{eq:sigmaEps}
\begin{split}
\sigma_\varepsilon^2(k,\theta) 
  &=  |A(k,\theta)|^2\sigma_Y^2(k) 
   +  |B(k,\theta)|^2\sigma_U^2(k) \\
  &- 2\RealPart \left( A(k,\theta) \overline{B(k,\theta)} \sigma_{YU}(k) \right)
\end{split}
\end{equation}
where $\sigma_Y^2$ is independent of $k$ for white noise, and the input error variance is assumed to be zero, i.e. $\sigma^2_U=0$, and similarly, the covariance $\sigma_{YU} = 0$.
Here, $\overline{B}$ denotes the complex conjugate of $B$.

@@

\begin{figure}[p]
  \setlength{\figurewidth}{0.85\columnwidth}
  \setlength{\figureheight}{0.68\figurewidth}
  \centering
  \ref{leg:init:stress:bodes}
  \input{\thisDir/fig/bodeplots.tikz}
  %TODO: put legend inside figure (note: subplot magic!!)
  \caption[Bode plots of the stress test filters.]{Bode plots of the tested filters~\legref{leg:bode:true} in the stress test. 
  For each filter, two (randomly selected) instances of the parametric estimates $\model{\bullet}$ are shown per method.
  % Top to bottom: Chebyshev I, Chebyshev II, Elliptical and Butterworth characteristic.
  % Left to right: low-pass, band-pass, high-pass and band-stop.
  }
  \label{fig:bodeplots}
\end{figure}

\begin{figure}[p]
  \setlength{\figurewidth}{0.85\columnwidth}
  \setlength{\figureheight}{0.68\figurewidth}
  \centering
  \ref{leg:distances}
  \input{\thisDir/fig/distances.tikz}
  \caption[Empirical \glsentrytext{CDF} of $\validationDistance{\bullet}$ in the stress test.]{Empirical cumulative distribution function of the validation distance $D_{\bullet} = \norm{\model{\bullet} - \true{G}}$ for the different tested systems (different numbered plots) and the different initial values (colors).
  Note that since $\norm{\true{G}}=1$ is chosen, $D_{\bullet}$ immediately indicates the relative RMS error of the estimates.
  }
  %TODO: put legend inside figure (note: subplot magic!!)
  \label{fig:distancesStress}
\end{figure}

@@

\begin{guideline}[Try regularization to improve convergence]
\label{guide:initvals:try-regularization}
Regularized estimators, such as \gls{RFIR}, sometimes provide better estimates than the `true' model parameters.
\end{guideline}

\begin{guideline}[Try different initial values for non-convex optimization]
\label{guide:initvals:try-many-initial-values}
Trying different initial values is quite inexpensive, but the quality of the final model may improve by several order of magnitude if a local minimum can be avoided.
As such, using multiple initial values, especially ones that have been derived with good judgment, pays off.
\end{guideline}

>>>>

In this chapter we assume that the output signal is disturbed by random noise, resulting in noisy error-prone data.
For the simulations, white noise is used ($\true{H} = 1$ and $e(t) = v(t)$).
It is also possible to apply the estimation procedure in this chapter to a system that is disturbed by colored noise, as will be demonstrated in \secref{se:ExpMeas}.

@@

With reference to equation \eqref{lpmtd1}, the relation between the noiseless input and the output signals ($v(t)= 0$) is assumed to be of the form
\begin{equation}
    A(q^{-1}) y_0(t) = B(q^{-1}) u_0(t) \quad \forall t \in n \Ts{}  \text{ with }  n \in \IntegerNumbers
\label{ABpolys}
\end{equation}
where $A$ and $B$ are polynomials in $q^{-1}$. 
Thus, it follows from equation \eqref{lpmtd1} that
\begin{equation}
    \true{G}(q^{-1})= \frac{B(q^{-1})}{A(q^{-1})}
    \text{.}
\end{equation}

From~\citet[Section 6.3.2.]{Pintelon2012}, in the frequency domain, and for the \glspl{DFT} of the windowed signals, equation \eqref{ABpolys} is of the form
\begin{align}
A(e^{-j\omega_k})\true{Y}(k) = B(e^{-j\omega_k})\true{U}(k) + I(e^{-j\omega_k})
\label{DFTspectra}
\end{align}
with $q^{-1}$ sampled in  $q_k^{-1} = e^{-j\omega_k}$, where $\omega_k = \frac{2\pi k}{N}$ are the \gls{DFT} frequencies, and $I$ is a polynomial of order $\max(N_a,N_b) - 1$, which depends on the difference between the initial and end conditions of the filters $\true{G}$ and $\true{H}$ during the measurement record.

@@

The maximum likelihood estimate of the parameter vector $\theta$  containing the coefficients of the $A$, $B$ and $I$  polynomials is obtained by solving the optimization problem
\begin{equation}
  \hat{\theta} = \arg\min_\theta V(\theta) \text{.}
\end{equation}
Since the noise $v(t)$ is assumed to be Gaussian, $V(\theta)$ accords to the weighted least squares cost function~\citep[Section 9.11]{Pintelon2012}:
\begin{equation}\label{eq:MLEcf}
V(\theta) = \sum_{k=1}^{N/2-1}\frac{|\varepsilon(k,\theta)|^2}{\sigma_\varepsilon^2(k,\theta)}
\end{equation}
for $N$ measurements and when $\varepsilon$ denotes the error in equation \eqref{DFTspectra},  \emph{viz}.:
\begin{align}
\varepsilon(k,\theta) = A(k,\theta)Y(k) - B(k,\theta)U(k) - I(k,\theta)\text{.}
\end{align}
The variance of the error is of the form:
\begin{equation}\label{eq:sigmaEps}
\begin{split}
\sigma_\varepsilon^2(k,\theta) 
  &=  |A(k,\theta)|^2\sigma_Y^2(k) 
   +  |B(k,\theta)|^2\sigma_U^2(k)
  - 2\RealPart \left( A(k,\theta) \overline{B(k,\theta)} \sigma_{YU}(k) \right)
\end{split}
\end{equation}
where $\sigma_Y^2$ is independent of $k$ for white noise, and the input error variance is assumed to be zero, i.e. $\sigma^2_U=0$, and similarly, the covariance $\sigma_{YU} = 0$.
Here, $\overline{B}$ denotes the complex conjugate of $B$.

@@

\begin{figure}[p]
  \setlength{\figurewidth}{0.85\columnwidth}
  \setlength{\figureheight}{0.68\figurewidth}
  \centering
  \input{\thisDir/fig/bodeplots.tikz}
  \caption[Bode plots of the stress test filters.]{Bode plots of the tested filters~\legref{leg:bode:true} in the stress test. 
  For each filter, two (randomly selected) instances of the parametric estimates $\model{\bullet}$ are shown per method.
  % Top to bottom: Chebyshev I, Chebyshev II, Elliptical and Butterworth characteristic.
  % Left to right: low-pass, band-pass, high-pass and band-stop.
  }
  \label{fig:bodeplots}
\end{figure}

\begin{figure}[p]
  \setlength{\figurewidth}{0.85\columnwidth}
  \setlength{\figureheight}{0.68\figurewidth}
  \centering
  \input{\thisDir/fig/distances.tikz}
  \caption[Empirical \glsentrytext{CDF} of $\validationDistance{\bullet}$ in the stress test.]{Empirical cumulative distribution function of the validation distance $D_{\bullet} = \norm{\model{\bullet} - \true{G}}$ for the different tested systems (different numbered plots) and the different initial values (colors).
  Note that since $\norm{\true{G}}=1$ is chosen, $D_{\bullet}$ immediately indicates the relative RMS error of the estimates.
  }
  \label{fig:distancesStress}
\end{figure}

@@

\begin{guideline}[Try regularization to improve convergence]
\label{guide:initvals:try-regularization}
Regularized estimators, such as \gls{RFIR}, sometimes provide better initial estimates than the `true' model parameters.
\end{guideline}

\begin{guideline}[Try different initial values for non-convex optimization]
\label{guide:initvals:try-many-initial-values}
Trying different initial values is quite inexpensive, but the quality of the final model may improve by several orders of magnitude if a local minimum can be avoided.
As such, using multiple initial values, especially ones that have been derived with good judgment, pays off.
\end{guideline}
